---
title: "World Development Indicators - Data Analysis"
author: "Sahil Verma"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction

This project analyzes World Development Indicators data sourced from the World Bank. The dataset covers **5 key indicators** across multiple countries from **2003 to 2022**:

1. GDP Growth (annual %)
2. Wage and Salaried Workers (%)
3. Machinery & Transport Equipment (% of manufacturing value added)
4. Fertility Rate (births per woman)
5. HepB3 Immunization (% of 1-year-olds)

---

## Load Libraries

```{r libraries}
library(tidyverse)
library(readxl)
library(wooldridge)
library(openxlsx)
```

---

## Import Data

```{r import}
# Update path if needed
sahil_verma <- read_excel("sahil verma.xlsx")
```

---

## Explore Data

```{r explore}
# Column names
names(sahil_verma)

# Summary statistics
summary(sahil_verma)
```

The dataset contains **100 rows** with data across multiple countries and years.

---

## Line Plots — Indicators Over Time

### 1. GDP Growth (annual %)

```{r gdp-growth}
ggplot(data = sahil_verma) +
  geom_smooth(mapping = aes(x = Time, y = GDP_growth_annual_percentage, color = Country_Name)) +
  labs(title = "GDP Growth (annual %) Over Time",
       x = "Year", y = "GDP Growth (%)", color = "Country") +
  theme_minimal()
```

### 2. Wage and Salaried Workers (%)

```{r wage}
ggplot(data = sahil_verma) +
  geom_smooth(mapping = aes(x = Time, y = wage_and_salaried_workers, color = Country_Name)) +
  labs(title = "Wage and Salaried Workers Over Time",
       x = "Year", y = "Wage Workers (%)", color = "Country") +
  theme_minimal()
```

### 3. Machinery & Transport Equipment (% of manufacturing)

```{r machinery}
ggplot(data = sahil_verma) +
  geom_smooth(mapping = aes(x = Time, y = Machinery_and_transport_equipment_percentage_of_value_added_in_manufacturing, color = Country_Name)) +
  labs(title = "Machinery & Transport Equipment Over Time",
       x = "Year", y = "Machinery & Transport (%)", color = "Country") +
  theme_minimal()
```

### 4. Fertility Rate (births per woman)

```{r fertility}
ggplot(data = sahil_verma) +
  geom_point(mapping = aes(x = Fertility_rate_total_births_per_woman, y = Time, color = Country_Name)) +
  labs(title = "Fertility Rate by Country",
       x = "Fertility Rate (births per woman)", y = "Year", color = "Country") +
  theme_minimal()
```

### 5. HepB3 Immunization (% of 1-year-olds)

```{r immunization}
ggplot(data = sahil_verma) +
  geom_smooth(mapping = aes(x = Time, y = Immunization_HepB3_pertcentage_of_one_year_old_children, color = Country_Name)) +
  labs(title = "HepB3 Immunization Over Time",
       x = "Year", y = "Immunization (%)", color = "Country") +
  theme_minimal()
```

---

## Box Plots — Distribution by Year

### 1. GDP Growth

```{r box-gdp}
ggplot(data = sahil_verma, mapping = aes(x = Time, y = GDP_growth_annual_percentage, color = Country_Name)) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "GDP Growth Distribution by Year",
       x = "Year", y = "GDP Growth (%)", color = "Country") +
  theme_minimal()
```

### 2. Wage and Salaried Workers

```{r box-wage}
ggplot(data = sahil_verma, mapping = aes(x = Time, y = wage_and_salaried_workers, color = Country_Name)) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "Wage Workers Distribution by Year",
       x = "Year", y = "Wage Workers (%)", color = "Country") +
  theme_minimal()
```

### 3. Machinery & Transport Equipment

```{r box-machinery}
ggplot(data = sahil_verma, mapping = aes(x = Time, y = Machinery_and_transport_equipment_percentage_of_value_added_in_manufacturing, color = Country_Name)) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "Machinery & Transport Distribution by Year",
       x = "Year", y = "Machinery & Transport (%)", color = "Country") +
  theme_minimal()
```

### 4. Fertility Rate

```{r box-fertility}
ggplot(data = sahil_verma, mapping = aes(x = Time, y = Fertility_rate_total_births_per_woman, color = Country_Name)) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "Fertility Rate Distribution by Year",
       x = "Year", y = "Fertility Rate (births per woman)", color = "Country") +
  theme_minimal()
```

---

## Conclusion

This analysis explored 5 World Development Indicators across multiple countries over approximately 20 years (2003–2022). Key observations:

- **GDP Growth** varies significantly across countries, with visible dips around global economic events.
- **Wage workers** show different trends depending on the country's development stage.
- **Fertility rates** have generally declined over time across most countries.
- **Immunization rates** show variation, with some countries having consistently lower coverage.

---

*Data Source: World Bank — World Development Indicators*
